Forecasting biodiversity in breeding birds using best practices

### 2017-11-20 manuscript/reviewer_reply.Rmd

*The reviewers and I agree that this manuscript is interesting, timely and well-written. I will add a few of my own observations here. First, and least important - normally submitted manuscripts do not hyphenate and do not right justify.*

*Next, in your figure 1, which forms the basis for comparison of model accuracy over time, I really expected you to choose data WITH a temporal pattern. After all, it looks to me like you have a time series in which species richness simply varies around a more or less stable average value, and the variation can be due to nothing (random), observer bias, and any number of species population trends (and combinations thereof). Thus, you have a situation of years of no trend, and use those data to predict no trend into the future? I would have thought you would choose data with a definite time trend that was already potentially correlated with, say, climate change, and then use hindcasting and forecasting to test whether the models accurately predict a trend into the "future"! To me, it is not surprising that models do not "...successfully turn spatial data into useful temporal predictions about biodiversity at decadal time-scales." But, I wonder if you chose data WITH a trend, you would have gotten other results.*

*Finally, your figure two seems to reverse panels A and B, neither of which is marked in the text.*

**We have switched the panels and labeled them. Thank you for noticing this error**.

@@

*Line 216: Cite randomForest / caret/ party R package – or whatever else you used?*

**The randomForest package was cited in the final methods paragraph. We've cited it where it is first mentioned as well.**

*Line 242: Again, I think including the equation here would be useful*

**We have added an in-line equation and clarified the wording**

@@

#### Methods:

*the authors perform their analyses on data that are collected from an incredibly dynamic time period. I wouldn’t expect species to have settled into a richness pattern given the incredibly disrupted nature of the last 32 years. Some strategies to try to handle these issues might be to train and predict at different temporal scales and to both forecast and hindcast the predictions. (e.g. train on 2004-2013 and try to predict 1982-1992 data + 1993-2003 data). If the authors were to make splits based on expected periods of dynamism, they could better tell whether the drivers that are included in the models are accurate predictors.*

*124-130: I didn’t see where future projections were performed/used in this paper. Why is this here?*

**Thank you for pointing out that we were unclear here.  The methods and purpose for the future predictions are now explained in more detail on lines [[XX]].**

### 2017-12-04 manuscript/reviewer_reply.Rmd

*Line 138: Maybe also cite any of the interfaces you used, like RStan: http://mc-stan.org/users/citations/*

**RStan is cited in the last paragraph of the Methods.**

*Line 140: I’m curious why you didn’t consider spatial structure in the random effects here, using GMRF models or similar? One hand these models might be overkill if the habitat/environmental variables (which are spatially correlated) explain variability in richness, but the spatial correlation may pick up other sources of variability not accounted for by those variables.*

**We agree that accounting for spatial autocorrelation in the observer model would have improved our estimates of the site-level random effects. However, we do not believe that this change would affect the paper's main results. Improvements would primarily benefit the "average" baseline, which is based directly on the observer model. As a result, it would be even harder for the three environmental models to outperform the baseline---especially since (as you noted), spatial and environmental explanations for variation in species richness trade off with one another.**

### 2017-12-06 manuscript/draft.Rmd
law of total variance
minor

---
title: "Forecasting biodiversity in breeding birds using best practices"
author:
- David J. Harris$^1$ (corresponding author)
- Shawn D. Taylor$^2$
- Ethan P. White$^1$
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \usepackage{booktabs}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: 
    keep_tex: true
  html_document: default
---

@@

We defined model performance for all models in terms of continuous Gaussian
errors, instead of using discrete count distributions. Variance in species
richness within sites was lower than predicted by several common count models,
such as the Poisson or binomial (i.e. richness was underdispersed for individual
sites), so these count models would have had difficulty fitting the data [cf.
@calabrese2014]. The use of a continuous distribution is adequate here, since
richness had a relatively large mean (`r numbers$richness_summary$mean`) and all
models produce continuous richness estimates. When a model was run multiple
times for the purpose of correcting for observer effects, we used the mean of
those runs' point estimates as our final point estimate and we calculated the
uncertainty using the law of total variance (i.e. the average of the model runs'
variance, plus the variance in the point estimates).

@@

![Partitioning of the squared error for each model into site and year
components. The site-level mean component shows consistent over or under
estimates of richness at a site across years. The annual fluctuation compoonent
shows errors in predicting fluctuations in a site's richness over time. Both
components of the mean squared error were lower for the single-site models than
for the environmental models.](../figures/barcharts.png)

>>>>

---
title: "Forecasting biodiversity in breeding birds using best practices"
author:
- David J. Harris$^1$ (corresponding author)
- Shawn D. Taylor$^2$
- Ethan P. White$^1$
bibliography: refs.bib
csl: csl.csl
geometry: left=3.5cm, right=3.5cm, top=2.25cm, bottom=2.25cm, headheight=12pt, letterpaper
fontsize: 12pt
header-includes:
    - \usepackage{times}
    - \usepackage{setspace}
    - \usepackage{booktabs}
    - \doublespacing
    - \usepackage{lineno}
    - \linenumbers
output:
  pdf_document: 
    keep_tex: true
  html_document: default
---
\raggedright

@@

We defined model performance for all models in terms of continuous Gaussian
errors, instead of using discrete count distributions. Variance in species
richness within sites was lower than predicted by several common count models,
such as the Poisson or binomial (i.e. richness was underdispersed for individual
sites), so these count models would have had difficulty fitting the data [cf.
@calabrese2014]. The use of a continuous distribution is adequate here, since
richness had a relatively large mean (`r numbers$richness_summary$mean`) and all
models produce continuous richness estimates. When a model was run multiple
times for the purpose of correcting for observer effects, we used the mean of
those runs' point estimates as our final point estimate and we calculated the
uncertainty using the law of total variance (i.e. $\mathrm{Var}(\bar{y}) + \mathop{{}\mathbb{E}}\big[\mathrm{Var}(y)\big]$, or the variance in point estimates plus the average residual variance).

@@

![Partitioning of the squared error for each model into site and year
components. The site-level mean component shows consistent over or under
estimates of richness at a site across years. The annual fluctuation component
shows errors in predicting fluctuations in a site's richness over time. Both
components of the mean squared error were lower for the single-site models than
for the environmental models.](../figures/barcharts.png)
